Parsing Postfix log files

### 2009-02-23 doc/logparser-introduction.tex
Annotated all URLs with last checked date.

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check if new anti-spam measures are working.  The more
adventurous will have generated statistics to show how successful each of
their anti-spam measures has been in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this kind of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from \url{http://www.postfix.org/}} Further
complication arises because the set of rejection messages is not fixed: new
messages can be added by the administrator with custom checks; every
\acronym{DNSBL}\footnote{This document is supplied with a glossary, see
\textsection\ref{Glossary}.} returns a different explanatory message;
policy servers may log different messages depending on the characteristics
of the connection.  There are many ways in which the log lines may differ
between servers, even within the same organisation: servers may be
configured differently, or running different versions of Postfix.  This
paper documents the difficult process of parsing Postfix log files,
presenting \parsername{}, a program that parses Postfix log files and
places the resulting data into a database for later use.  The gathered data
can then be used to optimise current anti-spam measures, provide a baseline
to test new anti-spam measures against, or to produce statistics showing
how effective those measures are.  Numerous other uses are possible for
such data: improving server performance by identifying troublesome
destinations and reconfiguring appropriately; identifying regular high
volume uses (e.g.\ customer newsletters) and restricting those uses to
off-peak times; detecting virus outbreaks that propagate via mail; as a
base for billing customers on a shared server.  Preserving the raw data
enables users to develop a multitude of uses far beyond those conceived of
by the author.

>>>>

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check if new anti-spam measures are working.  The more
adventurous will have generated statistics to show how successful each of
their anti-spam measures has been in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the author's first real foray into processing
Postfix log files.}  Very few will have performed in-depth parsing and
analysis of their log files, where the parsing must correlate the log lines
per-connection or per-queueid rather than processing log lines
independently.  One of the barriers to this kind of processing is the
unstructured nature of Postfix log files, where each log line was added on
an ad hoc basis as a requirement was discovered or new functionality was
added.\footnote{A history of all changes made to Postfix is distributed
with the source code, available from
\urlLastChecked{http://www.postfix.org/}{2009/02/23}} Further complication
arises because the set of rejection messages is not fixed: new messages can
be added by the administrator with custom checks; every
\acronym{DNSBL}\footnote{This document is supplied with a glossary, see
\textsection\ref{Glossary}.} returns a different explanatory message;
policy servers may log different messages depending on the characteristics
of the connection.  There are many ways in which the log lines may differ
between servers, even within the same organisation: servers may be
configured differently, or running different versions of Postfix.  This
paper documents the difficult process of parsing Postfix log files,
presenting \parsername{}, a program that parses Postfix log files and
places the resulting data into a database for later use.  The gathered data
can then be used to optimise current anti-spam measures, provide a baseline
to test new anti-spam measures against, or to produce statistics showing
how effective those measures are.  Numerous other uses are possible for
such data: improving server performance by identifying troublesome
destinations and reconfiguring appropriately; identifying regular high
volume uses (e.g.\ customer newsletters) and restricting those uses to
off-peak times; detecting virus outbreaks that propagate via mail; as a
base for billing customers on a shared server.  Preserving the raw data
enables users to develop a multitude of uses far beyond those conceived of
by the author.

### 2009-02-23 doc/logparser-other-parsers-review.tex
Annotated all URLs with last checked date.

There only appears to be one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \acronym{LMA} is
quite different from \parsername{}: it attempts to present correlated data
from log files in a form suitable for a systems administrator to search
using the myriad of standard Unix text processing utilities already
available.  It produces a \acronym{CSV} file and either a MySQL
(\url{http://www.mysql.com/}) or Berkeley DB
(\url{http://www.oracle.com/database/berkeley-db/index.html}) database.
The decision to support both \acronym{CSV} and Berkeley DB appears to have
been a serious limitation: both formats have limitations that will be
explained later.  Hardly any documentation is provided with \acronym{LMA},
though some documentation is available in~\cite{log-mail-analyser}.
Studying the source code is informative, though this author had difficulty
as the authors of \acronym{LMA} wrote in Italian.

@@

    \item \url{http://www.thesawmill.co.uk/formats/postfix.html} \newline{}
        Fields extracted: from, to, server, UID, relay, status, number of
        recipients, origin hostname, origin \acronym{IP} address, and
        virus.  It also counts the number of and total size of all mails
        delivered.  The fields \texttt{server}, \texttt{uid},
        \texttt{relay}, and \texttt{virus} are not explained in the
        documentation: \texttt{server} is probably the hostname or
        \acronym{IP} address of the server the mail is delivered to;
        \texttt{relay} might be the delivery method: \acronym{SMTP}, local
        delivery, or \acronym{LMTP}; \texttt{uid} might be the uid of the
        user submitting mail locally.  Postfix does not do any form of
        virus checking (though it has many options for cooperating with an
        external virus scanner), so the \texttt{virus} field is a mystery.

    \item \url{http://www.thesawmill.co.uk/formats/postfix_ii.html}
        \newline{} Fields extracted: from, to, RBL list, client hostname,
        and client \acronym{IP} address.  It also counts the number and
        total size of all mails delivered.

    \item \url{http://www.thesawmill.co.uk/formats/beta_postfix.html}
        \newline{} Fields extracted: from, to, client hostname, client
        \acronym{IP} address, relay hostname, relay \acronym{IP} address,
        status, response code, RBL list, and message id.  It also counts
        the number and size of all mails delivered, processed, blocked,
        expired, and bounced.

@@

Splunk aims to index all of an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface provides search functionality, generating statistics and
graphs in real time, a facility not provided by \parsername{}.  Splunk
allows quite complicated searches, based on the fields extracted by Splunk
(described later) or the full text of the log line, though it is not
possible to search on partial words.  Search results are not available for
use with other tools.  Searches can be saved for reuse; saved searches can
be run periodically and the results mailed to a recipient or sent to an
external program for further processing.  The author was unable to save
searches, though that may have been because of limitations in the free
version.  The data store is not available for use by external programs,
whereas \parsername{} provides the database and leaves it to the user to
utilise it without limit or restriction.  The web interface is optimised
for interactive use rather than automated queries, and it does not appear
to be possible to write independent tools to utilise the Splunk database.
Some additional Postfix reports are supposedly available on
\url{http://www.splunkbase.com/}, but the author was unable to find any
Postfix reports, or indeed reports for any other log file types: every
category was empty, even those that the web site claimed had numerous
reports available.  Many types of graphs can be generated, though most are
variations of a bar or pie chart, except the bubble and heatmap graphs.  It
is easy to drill down through the graphs to select a portion of the data
(e.g.\ select the hour with the largest number of events, then select a
particular host, and finally a specific sender address).  All searches
performed using the indexed data returned reasonable results.

@@

Copious documentation is made available on \url{http://www.splunk.com/},
but poor organisation and sheer abundance makes it extremely hard to find
useful information.  Searches confusingly tended to return results from old
documentation rather than new.  In general the documentation appears to
have been written by someone intimately acquainted with the software, who
has difficulty understanding how a newcomer would approach tasks or the
questions they would ask.

@@

AWStats can produce simple graphs from many different services' log files,
but supporting numerous log files formats without special purpose code
limits its functionality.  The data it can extract from an \acronym{MTA}
log file is limited in comparison to \parsername{}: time2, email, email\_r,
host, host\_r, method, url, code, and bytesd.  No explanation for any of
those fields is provided in the documentation at
\url{http://awstats.sourceforge.net/docs/awstats_faq.html#MAIL} (in
contrast \parsername{} fully documents everything it extracts), so the
author could not understand what data is extracted, or determine what data
is missing in comparison to \parsername{}.  AWStats coerces Postfix log
files into Apache\footnote{The Apache web server is the most popular HTTP
server in use over the past 10 years; more information is available at
\url{http://httpd.apache.org/}.} format log files, for analysis by AWStats'
HTTP log file parser.  The converting parser only deals with a small
portion of the log lines generated by Postfix, silently skipping those it
cannot deal with, and does not distinguish between different types of
rejection; extending it to parse all log lines would be at least as much
work as writing a new parser.  It does correlate log lines by queueid (not
by pid), but it does not deal with any of the other complications described
in \sectionref{complications}.  AWStats supports saving data, but the
format of the data store is not documented.  It also supports reading
compressed log files, but that functionality was not tested.

@@

The Debian project (\url{http://www.debian.org/}) provides a manual page
with the copy of anteater it distributes, so the author was at least able
to run anteater with the correct arguments; sadly anteater produced zero
for every statistic, presumably because it was unsuccessful in parsing the
log lines.

@@

YAALA provides a mini-language based on \acronym{SQL} that is used when
generating reports; sample reports can be seen
at~\url{http://www.yaala.org/samples.html}.  Example query for HTTP proxy
servers: \newline{} \tab{} \texttt{requests BY file WHERE host =\~{}
Google} \newline{} The mini-language is quite limited and cannot be used to
extract data for external use, merely to create reports.  Only data
selected by the query will be saved in the data store; other data will be
discarded, and removed from the data store if already present.

>>>>

There only appears to be one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \acronym{LMA} is
quite different from \parsername{}: it attempts to present correlated data
from log files in a form suitable for a systems administrator to search
using the myriad of standard Unix text processing utilities already
available.  It produces a \acronym{CSV} file and either a MySQL
(\urlLastChecked{http://www.mysql.com/}{2008/11/23}) or Berkeley DB
(\urlLastChecked{http://www.oracle.com/database/berkeley-db/index.html}{2008/11/23})
database.  The decision to support both \acronym{CSV} and Berkeley DB
appears to have been a serious limitation: both formats have limitations
that will be explained later.  Hardly any documentation is provided with
\acronym{LMA}, though some documentation is available
in~\cite{log-mail-analyser}.  Studying the source code is informative,
though this author had difficulty as the authors of \acronym{LMA} wrote in
Italian.

@@

    \item
        \urlLastChecked{http://www.thesawmill.co.uk/formats/postfix.html}{2008/11/23}
        \newline{} Fields extracted: from, to, server, UID, relay, status,
        number of recipients, origin hostname, origin \acronym{IP} address,
        and virus.  It also counts the number of and total size of all
        mails delivered.  The fields \texttt{server}, \texttt{uid},
        \texttt{relay}, and \texttt{virus} are not explained in the
        documentation: \texttt{server} is probably the hostname or
        \acronym{IP} address of the server the mail is delivered to;
        \texttt{relay} might be the delivery method: \acronym{SMTP}, local
        delivery, or \acronym{LMTP}; \texttt{uid} might be the uid of the
        user submitting mail locally.  Postfix does not do any form of
        virus checking (though it has many options for cooperating with an
        external virus scanner), so the \texttt{virus} field is a mystery.

    \item
        \urlLastChecked{http://www.thesawmill.co.uk/formats/postfix_ii.html}{2008/11/23}
        \newline{} Fields extracted: from, to, RBL list, client hostname,
        and client \acronym{IP} address.  It also counts the number and
        total size of all mails delivered.

    \item
        \urlLastChecked{http://www.thesawmill.co.uk/formats/beta_postfix.html}{2008/11/23}
        \newline{} Fields extracted: from, to, client hostname, client
        \acronym{IP} address, relay hostname, relay \acronym{IP} address,
        status, response code, RBL list, and message id.  It also counts
        the number and size of all mails delivered, processed, blocked,
        expired, and bounced.

@@

Splunk aims to index all of an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface provides search functionality, generating statistics and
graphs in real time, a facility not provided by \parsername{}.  Splunk
allows quite complicated searches, based on the fields extracted by Splunk
(described later) or the full text of the log line, though it is not
possible to search on partial words.  Search results are not available for
use with other tools.  Searches can be saved for reuse; saved searches can
be run periodically and the results mailed to a recipient or sent to an
external program for further processing.  The author was unable to save
searches, though that may have been because of limitations in the free
version.  The data store is not available for use by external programs,
whereas \parsername{} provides the database and leaves it to the user to
utilise it without limit or restriction.  The web interface is optimised
for interactive use rather than automated queries, and it does not appear
to be possible to write independent tools to utilise the Splunk database.
Some additional Postfix reports are supposedly available on
\urlLastChecked{http://www.splunkbase.com/}{2008/11/23}, but the author was
unable to find any Postfix reports, or indeed reports for any other log
file types: every category was empty, even those that the web site claimed
had numerous reports available.  Many types of graphs can be generated,
though most are variations of a bar or pie chart, except the bubble and
heatmap graphs.  It is easy to drill down through the graphs to select a
portion of the data (e.g.\ select the hour with the largest number of
events, then select a particular host, and finally a specific sender
address).  All searches performed using the indexed data returned
reasonable results.

@@

Copious documentation is made available on
\urlLastChecked{http://www.splunk.com/}{2008/11/23}, but poor organisation
and sheer abundance makes it extremely hard to find useful information.
Searches confusingly tended to return results from old documentation rather
than new.  In general the documentation appears to have been written by
someone intimately acquainted with the software, who has difficulty
understanding how a newcomer would approach tasks or the questions they
would ask.

@@

AWStats can produce simple graphs from many different services' log files,
but supporting numerous log files formats without special purpose code
limits its functionality.  The data it can extract from an \acronym{MTA}
log file is limited in comparison to \parsername{}: time2, email, email\_r,
host, host\_r, method, url, code, and bytesd.  No explanation for any of
those fields is provided in the documentation at
\urlLastChecked{http://awstats.sourceforge.net/docs/awstats_faq.html\#MAIL}{2008/11/23}
(in contrast \parsername{} fully documents everything it extracts), so the
author could not understand what data is extracted, or determine what data
is missing in comparison to \parsername{}.  AWStats coerces Postfix log
files into Apache\footnote{The Apache web server is the most popular HTTP
server in use over the past 10 years; more information is available at
\urlLastChecked{http://httpd.apache.org/}{2008/11/23}.} format log files,
for analysis by AWStats' HTTP log file parser.  The converting parser only
deals with a small portion of the log lines generated by Postfix, silently
skipping those it cannot deal with, and does not distinguish between
different types of rejection; extending it to parse all log lines would be
at least as much work as writing a new parser.  It does correlate log lines
by queueid (not by pid), but it does not deal with any of the other
complications described in \sectionref{complications}.  AWStats supports
saving data, but the format of the data store is not documented.  It also
supports reading compressed log files, but that functionality was not
tested.

@@

The Debian project (\urlLastChecked{http://www.debian.org/}{2008/11/23})
provides a manual page with the copy of anteater it distributes, so the
author was at least able to run anteater with the correct arguments; sadly
anteater produced zero for every statistic, presumably because it was
unsuccessful in parsing the log lines.

@@

YAALA provides a mini-language based on \acronym{SQL} that is used when
generating reports; sample reports can be seen
at~\urlLastChecked{http://www.yaala.org/samples.html}{2008/11/23}.  Example
query for HTTP proxy servers: \newline{} \tab{} \texttt{requests BY file
WHERE host =\~{} Google} \newline{} The mini-language is quite limited and
cannot be used to extract data for external use, merely to create reports.
Only data selected by the query will be saved in the data store; other data
will be discarded, and removed from the data store if already present.

### 2009-02-23 doc/logparser-postfix-daemons.tex
Annotated all URLs with last checked date.

% vim: set filetype=tex :
% Postfix components
\newglossaryentry{bounce}{name={bounce},type={postfix},description={
    The bounce daemon is responsible for sending bounce notifications in
    Postfix versions later than 2.2.  The definitive documentation is
    \url{http://www.postfix.org/bounce.8.html}.
}}

\newglossaryentry{cleanup}{name={cleanup},type={postfix},description={
    Cleanup processes all incoming mail after it has been accepted and
    before it is delivered.  It removes duplicate recipient addresses,
    inserts missing headers, and optionally rewrites addresses if
    configured to do so.  The definitive documentation is
    \url{http://www.postfix.org/cleanup.8.html}.
}}

\newglossaryentry{lmtp}{name={lmtp},type={postfix},description={
    Delivery of mail over \acronym{LMTP} is performed by the lmtp
    component.  The definitive documentation is
    \url{http://www.postfix.org/lmtp.8.html}.
}}

\newglossaryentry{local}{name={local},type={postfix},description={
    Local is the Postfix component responsible for local delivery of mail
    (i.e.\ delivered on the server Postfix is running on), whether it be to
    a user's mailbox or a program such as a mailing list manager or
    procmail (\url{http://www.procmail.org/}).  It also handles aliases and
    processing of a user's \texttt{.forward} file.  The definitive
    documentation is \url{http://www.postfix.org/local.8.html}.
}}

\newglossaryentry{pickup}{name={pickup},type={postfix},description={
    Pickup is the service that deals with mail submitted locally via
    postdrop and sendmail; it passes all submitted mail to cleanup.  The
    definitive documentation is \url{http://www.postfix.org/pickup.8.html}.
}}

\newglossaryentry{postdrop}{name={postdrop},type={postfix},description={
    Postdrop is used when submitting mail locally on the server: it creates
    a new mail in the queue and copies its input into the mail.  Subsequent
    delivery of the mail is the responsibility of other Postfix components.
    The definitive documentation is
    \url{http://www.postfix.org/postdrop.1.html}.
}}

\newglossaryentry{postsuper}{name={postsuper},type={postfix},description={
    Maintenance tasks such as deleting mails from the queue, putting mail
    on hold and later releasing it (no further delivery attempts will be
    made until it is released), and consistency checking of the mail queue.
    The definitive documentation is available at
    \url{http://www.postfix.org/postsuper.1.html}.
}}

\newglossaryentry{qmgr}{name={qmgr},type={postfix},description={
    Qmgr is the Postfix daemon that manages the mail queue, determining
    which mails will be delivered next.  Qmgr orders the mails based on the
    recipient for local mails and the destination server for remote
    addresses, ensuring that it balances the aims of achieving maximum
    concurrency while avoiding overwhelming destinations or wasting time
    and resources on non-responsive destinations.  The definitive
    documentation is \url{http://www.postfix.org/qmgr.8.html}.
}}

\newglossaryentry{sendmail}{name={sendmail},type={postfix},description={
    Postfix provides a command that is compatible with the Sendmail
    (\url{http://www.sendmail.org/}) mail submission program that all Unix
    commands that send mail depend on; it executes \daemon{postdrop} to
    place a new mail in the queue.  The definitive documentation is
    \url{http://www.postfix.org/sendmail.1.html}.
}}

\newglossaryentry{smtp}{name={smtp},type={postfix},description={
    Delivery of mail over \acronym{SMTP} is performed by the smtp
    component.  The definitive documentation is
    \url{http://www.postfix.org/smtp.8.html}.
}}

\newglossaryentry{smtpd}{name={smtpd},type={postfix},description={
    Smtpd is the Postfix program that accepts mail via \acronym{SMTP}, and
    implements all the anti-spam restrictions Postfix provides.  The
    definitive documentation is \url{http://www.postfix.org/smtpd.8.html}.
}}

\newglossaryentry{virtual}{name={virtual},type={postfix},description={
    Virtual is the Postfix component responsible for delivery of mails to
    virtual domains.  With \daemon{local} delivery the destination is
    determined only by the portion of the email address on the left side of
    the @, whereas with \daemon{virtual} delivery the destination is
    determined by the entire email address.  E.g.\ if the server is
    responsible for both the \textbf{example.org} and \textbf{example.net}
    domains: \daemon{local} considers \textbf{john@example.org} and
    \textbf{john@example.net} to have the same mailbox, whereas
    \daemon{virtual} considers them to have different mailboxes.  Virtual
    delivery is used when a server hosts multiple domains where a username
    may be present in more than one domain but represent different users in
    each.  The definitive documentation is
    \url{http://www.postfix.org/virtual.8.html}.
}}

>>>>

% vim: set filetype=tex :
% Postfix components
\newglossaryentry{bounce}{name={bounce},type={postfix},description={
    The bounce daemon is responsible for sending bounce notifications in
    Postfix versions later than 2.2.  The definitive documentation is
    \urlLastChecked{http://www.postfix.org/bounce.8.html}{2009/02/23}.
}}

\newglossaryentry{cleanup}{name={cleanup},type={postfix},description={
    Cleanup processes all incoming mail after it has been accepted and
    before it is delivered.  It removes duplicate recipient addresses,
    inserts missing headers, and optionally rewrites addresses if
    configured to do so.  The definitive documentation is
    \urlLastChecked{http://www.postfix.org/cleanup.8.html}{2009/02/23}.
}}

\newglossaryentry{lmtp}{name={lmtp},type={postfix},description={
    Delivery of mail over \acronym{LMTP} is performed by the lmtp
    component.  The definitive documentation is
    \urlLastChecked{http://www.postfix.org/lmtp.8.html}{2009/02/23}.
}}

\newglossaryentry{local}{name={local},type={postfix},description={
    Local is the Postfix component responsible for local delivery of mail
    (i.e.\ delivered on the server Postfix is running on), whether it be to
    a user's mailbox or a program such as a mailing list manager or
    procmail (\urlLastChecked{http://www.procmail.org/}{2009/02/23}).  It
    also handles aliases and processing of a user's \texttt{.forward} file.
    The definitive documentation is
    \urlLastChecked{http://www.postfix.org/local.8.html}{2009/02/23}.
}}

\newglossaryentry{pickup}{name={pickup},type={postfix},description={
    Pickup is the service that deals with mail submitted locally via
    postdrop and sendmail; it passes all submitted mail to cleanup.  The
    definitive documentation is
    \urlLastChecked{http://www.postfix.org/pickup.8.html}{2009/02/23}.
}}

\newglossaryentry{postdrop}{name={postdrop},type={postfix},description={
    Postdrop is used when submitting mail locally on the server: it creates
    a new mail in the queue and copies its input into the mail.  Subsequent
    delivery of the mail is the responsibility of other Postfix components.
    The definitive documentation is
    \urlLastChecked{http://www.postfix.org/postdrop.1.html}{2009/02/23}.
}}

\newglossaryentry{postsuper}{name={postsuper},type={postfix},description={
    Maintenance tasks such as deleting mails from the queue, putting mail
    on hold and later releasing it (no further delivery attempts will be
    made until it is released), and consistency checking of the mail queue.
    The definitive documentation is available at
    \urlLastChecked{http://www.postfix.org/postsuper.1.html}{2009/02/23}.
}}

\newglossaryentry{qmgr}{name={qmgr},type={postfix},description={
    Qmgr is the Postfix daemon that manages the mail queue, determining
    which mails will be delivered next.  Qmgr orders the mails based on the
    recipient for local mails and the destination server for remote
    addresses, ensuring that it balances the aims of achieving maximum
    concurrency while avoiding overwhelming destinations or wasting time
    and resources on non-responsive destinations.  The definitive
    documentation is
    \urlLastChecked{http://www.postfix.org/qmgr.8.html}{2009/02/23}.
}}

\newglossaryentry{sendmail}{name={sendmail},type={postfix},description={
    Postfix provides a command that is compatible with the Sendmail
    (\urlLastChecked{http://www.sendmail.org/}{2009/02/23}) mail submission
    program that all Unix commands that send mail depend on; it executes
    \daemon{postdrop} to place a new mail in the queue.  The definitive
    documentation is
    \urlLastChecked{http://www.postfix.org/sendmail.1.html}{2009/02/23}.
}}

\newglossaryentry{smtp}{name={smtp},type={postfix},description={
    Delivery of mail over \acronym{SMTP} is performed by the smtp
    component.  The definitive documentation is
    \urlLastChecked{http://www.postfix.org/smtp.8.html}{2009/02/23}.
}}

\newglossaryentry{smtpd}{name={smtpd},type={postfix},description={
    Smtpd is the Postfix program that accepts mail via \acronym{SMTP}, and
    implements all the anti-spam restrictions Postfix provides.  The
    definitive documentation is
    \urlLastChecked{http://www.postfix.org/smtpd.8.html}{2009/02/23}.
}}

\newglossaryentry{virtual}{name={virtual},type={postfix},description={
    Virtual is the Postfix component responsible for delivery of mails to
    virtual domains.  With \daemon{local} delivery the destination is
    determined only by the portion of the email address on the left side of
    the @, whereas with \daemon{virtual} delivery the destination is
    determined by the entire email address.  E.g.\ if the server is
    responsible for both the \textbf{example.org} and \textbf{example.net}
    domains: \daemon{local} considers \textbf{john@example.org} and
    \textbf{john@example.net} to have the same mailbox, whereas
    \daemon{virtual} considers them to have different mailboxes.  Virtual
    delivery is used when a server hosts multiple domains where a username
    may be present in more than one domain but represent different users in
    each.  The definitive documentation is
    \urlLastChecked{http://www.postfix.org/virtual.8.html}{2009/02/23}.
}}
